Accounting for uncertainty in simulation greenhouse gas emissions from dairy cattle in Kenya

### 2022-06-06 Manuscript_analyses.Rmd
diagnostic plots for breed and herd structure interventions

#install and load packages
# devtools::install_github("CWWhitney/uncertainty")
library(DiagrammeR)
library(decisionSupport)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(plyr)
library(rmarkdown)
library(uncertainty)
library(chillR)
library(raster)

@@

```{r, warning=FALSE, eval=FALSE}
devtools::install_github("CWWhitney/uncertainty")
library(DiagrammeR)
library(decisionSupport)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(plyr)
library(rmarkdown)
library(uncertainty)
library(chillR)
library(raster)
```

@@

histo + geom_vline(data=farm5_emissions_precise$y, aes(xintercept=CO2eq_per_milk),color="blue", lwd=3)

@@

ggplot(data=data.frame(CO2eq_per_milk=slice$resampled), aes(CO2eq_per_milk)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") +
    xlab(expression(Emission~intensity~of~milk~production~at~2000~kg~milk~per~capita~(kg~CO[2]*'-eq'~kg^-1)))

@@

ggplot(data=slices, aes(Values, fill=Level)) +
 geom_density(alpha=.2) +
  ylab("Relative probability") +
    xlab(expression(Emission~intensity~of~milk~production~at~specific~milk~yield~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Production level')

@@

ggplot(data = decreased_emissions_probability, aes(Base_yield, new_yield, z =
                                                      probability)) + geom_contour_filled() +
  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
       fill="Probability of \nlower emissions \nintensity") +
  stat_contour(breaks = c(0.9),col="black")

@@

plot_nobull <- ggplot(plotdata_breed, aes(x = in_var, y = out_var)) + 
  geom_point(shape = 20, color = "blue", size = 2) + 
  stat_smooth(method = "loess", fullrange = TRUE) +
  geom_density_2d() +
  geom_rug(color="lightskyblue3") + 
  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
  theme_bw() +
  theme(legend.position = c(0.15, 0.9)) + ggtitle("'Upgrading breeds' scenario")

@@

dens_nobull<-ggplot(plotdata_breed, aes(x = in_var, y = out_var)) +
  stat_density_2d(
    geom = "raster",
    aes(fill = after_stat(density)),
    contour = FALSE
    ) +
  scale_fill_viridis_c() + 
  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
  theme_bw() + guides(fill="none")

@@

@@

breed_res$y<-breed_res$y[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
breed_res$x<-breed_res$x[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]

@@

breed_probs<-ggplot(data = decreased_emissions_probability,
                    aes(Base_yield, new_yield)) +
  geom_contour_filled(aes(z=probability)) +
  stat_contour(breaks = c(0.9),
               aes(z=probability),col="black") +
  geom_point(data=impact_of_breed[which(impact_of_breed$milk_change>0),],aes(x=baseline,y=breed_milk)) +
  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
       fill="Probability of \nlower emissions \nintensity")

@@

@@

ggsave("Fig_13_no_bulls_scatter_per_head_milk_emissions_intensity.png",width=9,height=9)

@@

nobull_res$y<-nobull_res$y[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
nobull_res$x<-nobull_res$x[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]

@@

nobull_probs<-ggplot(data = decreased_emissions_probability,
                    aes(Base_yield, new_yield)) +
  geom_contour_filled(aes(z=probability)) +
  stat_contour(breaks = c(0.9),
               aes(z=probability),col="black") +
  geom_point(data=impact_of_nobull[which(impact_of_nobull$milk_change>0),],aes(x=baseline,y=nobull_milk)) +
  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
       fill="Probability of \nlower emissions \nintensity")

@@

ggsave("Fig_14_nobull_reductions.png", width=9,height=9)

@@

lessReplace_res$y<-lessReplace_res$y[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
lessReplace_res$x<-lessReplace_res$x[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]

@@

lessRep_probs<-ggplot(data = decreased_emissions_probability,
                      aes(Base_yield, new_yield)) +
  geom_contour_filled(aes(z=probability)) +
  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
  geom_point(data=impact_of_lessReplace[which(impact_of_lessReplace$milk_change>0),],
             aes(x=baseline,y=lessReplace_milk)) +
  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
       fill="Probability of \nlower emissions \nintensity")

@@

ggsave("Fig_15_lessReplace_reductions.png", width=9,height=9)

>>>>

#install and load packages
# devtools::install_github("CWWhitney/uncertainty")
library(DiagrammeR)
library(decisionSupport)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(plyr)
library(rmarkdown)
library(uncertainty)
library(chillR)
library(raster)
library(reshape2)

@@

```{r, warning=FALSE, eval=FALSE}
devtools::install_github("CWWhitney/uncertainty")
library(DiagrammeR)
library(decisionSupport)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(plyr)
library(rmarkdown)
library(uncertainty)
library(chillR)
library(raster)
library(reshape2)
```

@@

histo + geom_vline(data=farm5_emissions_precise$y, aes(xintercept=CO2eq_per_milk),color="blue", lwd=3) + theme_bw()

@@

ggplot(data=data.frame(CO2eq_per_milk=slice$resampled), aes(CO2eq_per_milk)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") +
    xlab(expression(Emission~intensity~of~milk~production~at~2000~kg~milk~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + theme_bw()

@@

ggplot(data=slices, aes(Values, fill=Level)) +
 geom_density(alpha=.2) +
  ylab("Relative probability") +
    xlab(expression(Emission~intensity~of~milk~production~at~specific~milk~yield~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Production level') + theme_bw()

@@

ggplot(data = decreased_emissions_probability, aes(Base_yield, new_yield, z =
                                                      probability)) + geom_contour_filled() +
  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
       fill="Probability of \nlower emissions \nintensity") +
  stat_contour(breaks = c(0.9),col="black") + theme_bw()

@@

plot_breed <- ggplot(plotdata_breed, aes(x = in_var, y = out_var)) + 
  geom_point(shape = 20, color = "blue", size = 2) + 
  stat_smooth(method = "loess", fullrange = TRUE) +
  geom_density_2d() +
  geom_rug(color="lightskyblue3") + 
  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
  theme_bw() +
  theme(legend.position = c(0.15, 0.9)) + ggtitle("'Upgrading breeds' scenario")

@@

dens_breed<-ggplot(plotdata_breed, aes(x = in_var, y = out_var)) +
  stat_density_2d(
    geom = "raster",
    aes(fill = after_stat(density)),
    contour = FALSE
    ) +
  scale_fill_viridis_c() + 
  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
  theme_bw() + guides(fill="none")

@@

plot_breed + plot_baseline + dens_breed + dens_baseline +
  plot_layout(ncol = 2, nrow = 2)

@@

breed_res$y<-breed_res$y[normal_milk_yield,]
breed_res$x<-breed_res$x[normal_milk_yield,]

@@

breed_probs<-ggplot(data = decreased_emissions_probability,
                    aes(Base_yield, new_yield)) +
  geom_contour_filled(aes(z=probability)) +
  stat_contour(breaks = c(0.9),
               aes(z=probability),col="black") +
  geom_point(data=impact_of_breed[which(impact_of_breed$milk_change>0),],aes(x=baseline,y=breed_milk)) +
  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
       fill="Probability of \nlower emissions \nintensity") + theme_bw()

@@

### Characteristics of sensitive farms

The figure above shows the probability of detecting changes in greenhouse gas emissions intensity by comparing milk yields, for farms that have implemented the breed upgrade intervention. We can now extract the most likely probability of such detection for each of the farms (based on the nearest data point on the probability surface). This allows us to detect farms that are associated with particularly high probabilities (e.g. 90%). We can then look in detail at the characteristics of these farms to identify farms that are particularly likely to benefit (in terms of reducing emissions intensities) from the breed upgrade intervention.

breed_diagnosis<-breed_res
breed_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield

# get closest probability value

for(i in 1:nrow(breed_diagnosis$x))
  {decreased_emissions_probability[,"diff"]<-
    abs(decreased_emissions_probability$Base_yield-breed_diagnosis$y$baseline_pc_milk_yield[i]) +
    abs(decreased_emissions_probability$new_yield-breed_diagnosis$y$pc_milk_yield[i])
  breed_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
}

breed_90_farms<-breed_diagnosis$x$Scenario[which(breed_diagnosis$y$confidence_level>0.9)]

The only farm characteristics that we modified in this intervention is the breed composition, so the success of the intervention can only depend on the array of dairy cow breeds present on the farms. We can plot the breed composition of the farms against the probability that higher per-capita milk production is associated with lower emissions intensity.

farm_structures<-data.frame(Farm=unique(Kenyaherds$QQNO))

for(f in farm_structures$Farm)
  {samp<-Kenyaherds[which(Kenyaherds$QQNO==f),]
   cows<-samp[which(samp$animal_type %in% 4:6),]
   farm_structures[which(farm_structures$Farm==f),"breed_1"]<-length(which(cows$breed==1))/nrow(cows)
   farm_structures[which(farm_structures$Farm==f),"breed_2"]<-length(which(cows$breed==2))/nrow(cows)
   farm_structures[which(farm_structures$Farm==f),"breed_3"]<-length(which(cows$breed==3))/nrow(cows)
   farm_structures[which(farm_structures$Farm==f),"breed_4"]<-length(which(cows$breed==4))/nrow(cows)
   farm_structures[which(farm_structures$Farm==f),"breed_5"]<-length(which(cows$breed==5))/nrow(cows)
   farm_structures[which(farm_structures$Farm==f),"breed_6"]<-length(which(cows$breed==6))/nrow(cows)
   farm_structures[which(farm_structures$Farm==f),"breed_7"]<-length(which(cows$breed==7))/nrow(cows)
   farm_structures[which(farm_structures$Farm==f),"breed_8"]<-length(which(cows$breed==8))/nrow(cows)
   farm_structures[which(farm_structures$Farm==f),"unproductive_bulls"]<-
     length(which(samp$animal_type==1))/nrow(samp)
   farm_structures[which(farm_structures$Farm==f),"replacement_males"]<-
     length(which(samp$animal_type==3))/nrow(samp)

breed_diagnosis$x["Farm"]<-sapply(breed_diagnosis$x$Scenario,
                                  function(x) as.numeric(strsplit(x,"_")[[1]][2]))

breed_structures<-merge(breed_diagnosis$x,farm_structures)
breed_structures<-list(x=breed_structures, y=breed_diagnosis$y)

to_plot<-cbind(Farm=breed_structures$x$Farm,Confidence_level=breed_structures$y$confidence_level,
               breed_structures$x[,c(paste0("breed_",1:8),"unproductive_bulls","replacement_males")])

melted<-melt(to_plot, id.vars=c("Farm","Confidence_level","unproductive_bulls","replacement_males"))
melted<-melted[which(!is.na(melted$value)),]

melted$Confidence_level_rounded<-round(melted$Confidence_level,2)*100

breedcol<-palette.colors(n = 8, palette = "Okabe-Ito", alpha=1, recycle = FALSE)

ggplot(melted, aes(group=variable, fill=variable, y=value, x=Confidence_level_rounded)) + 
    geom_bar(position="fill", stat="identity") + coord_flip() +
  labs(y="Cumulative share of breed in herd",
       x="Probability of lower emissions intensity (%)",
       fill="Breed of \ndairy cows") + scale_fill_manual(values= c("breed_1" = breedcol[1],
                                                                   "breed_2" = breedcol[2],
                                                                   "breed_3" = breedcol[3],
                                                                   "breed_4" = breedcol[4],
                                                                   "breed_5" = breedcol[5],
                                                                   "breed_6" = breedcol[6],
                                                                   "breed_7" = breedcol[7],
                                                                   "breed_8" = breedcol[8]),
                                               labels = c("Holstein Friesian (pure/improved)",
                                                          "Ayrshire (pure/improved)",
                                                          "Jersey (pure/improved)",
                                                          "Guernsey (pure/improved)",
                                                          "Cross-bred unknown",
                                                          "Sahiwal",
                                                          "Boran",
                                                          "Local zebu")) +
  theme_bw()

ggsave("Fig_13_breed_diagnostics.png",width=5,height=5)

@@

ggsave("Fig_14_no_bulls_scatter_per_head_milk_emissions_intensity.png",width=9,height=9)

@@

nobull_res$y<-nobull_res$y[normal_milk_yield,]
nobull_res$x<-nobull_res$x[normal_milk_yield,]

@@

nobull_probs<-ggplot(data = decreased_emissions_probability,
                    aes(Base_yield, new_yield)) +
  geom_contour_filled(aes(z=probability)) +
  stat_contour(breaks = c(0.9),
               aes(z=probability),col="black") +
  geom_point(data=impact_of_nobull[which(impact_of_nobull$milk_change>0),],aes(x=baseline,y=nobull_milk)) +
  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
       fill="Probability of \nlower emissions \nintensity") + theme_bw()

@@

ggsave("Fig_15_nobull_reductions.png", width=9,height=9)

The confidence level depends on the farms' herd structure, specifically on the share of unproductive males with the farms' herds.

nobull_diagnosis<-nobull_res
nobull_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield

# get closest probability value

for(i in 1:nrow(nobull_diagnosis$x))
  {decreased_emissions_probability[,"diff"]<-
    abs(decreased_emissions_probability$Base_yield-nobull_diagnosis$y$baseline_pc_milk_yield[i]) +
    abs(decreased_emissions_probability$new_yield-nobull_diagnosis$y$pc_milk_yield[i])
  nobull_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
}

nobull_90_farms<-nobull_diagnosis$x$Scenario[which(nobull_diagnosis$y$confidence_level>0.9)]

nobull_diagnosis$x["Farm"]<-sapply(nobull_diagnosis$x$Scenario,
                                  function(x) as.numeric(strsplit(x,"_")[[1]][2]))

nobull_structures<-merge(nobull_diagnosis$x,farm_structures)
nobull_structures<-list(x=nobull_structures, y=nobull_diagnosis$y)

to_plot<-cbind(Farm=nobull_structures$x$Farm,Confidence_level=nobull_structures$y$confidence_level,
               nobull_structures$x[,c(paste0("breed_",1:8),"unproductive_bulls","replacement_males")])
to_plot$other_animals<-1-to_plot$unproductive_bulls

melted<-melt(to_plot[,c("Farm","Confidence_level","other_animals","unproductive_bulls")], id.vars=c("Farm","Confidence_level"))
melted<-melted[which(!is.na(melted$value)),]

melted$Confidence_level_rounded<-round(to_plot$Confidence_level,2)*100

ggplot(melted, aes(group=variable, fill=variable, y=value , x=Confidence_level_rounded)) + 
    geom_bar(position="fill", stat="identity") + coord_flip() +
  labs(y="Share of animal type in the herd",
       x="Probability of lower emissions intensity (%)",
       fill="Animal type") + scale_fill_manual(values= c("unproductive_bulls" = "burlywood",
                                                         "other_animals" = "darkslategrey"),
                                               labels = c("Unproductive bulls","Other animals")) +
  theme_bw()

ggsave("Fig_16_nobull_diagnostics.png",width=5,height=5)

@@

lessReplace_res$y<-lessReplace_res$y[normal_milk_yield,]
lessReplace_res$x<-lessReplace_res$x[normal_milk_yield,]

@@

lessRep_probs<-ggplot(data = decreased_emissions_probability,
                      aes(Base_yield, new_yield)) +
  geom_contour_filled(aes(z=probability)) +
  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
  geom_point(data=impact_of_lessReplace[which(impact_of_lessReplace$milk_change>0),],
             aes(x=baseline,y=lessReplace_milk)) +
  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
       fill="Probability of \nlower emissions \nintensity") + theme_bw()

@@

ggsave("Fig_17_lessReplace_reductions.png", width=9,height=9)

Also with this intervention, the level of confidence we can have in higher per-capita milk yields being associated with lower emissions intensity is related to the farms' herd structures. In this case, the share of replacement males within the herds is a key determinant.

lessReplace_diagnosis<-lessReplace_res
lessReplace_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield

# get closest probability value

for(i in 1:nrow(lessReplace_diagnosis$x))
  {decreased_emissions_probability[,"diff"]<-
    abs(decreased_emissions_probability$Base_yield-lessReplace_diagnosis$y$baseline_pc_milk_yield[i]) +
    abs(decreased_emissions_probability$new_yield-lessReplace_diagnosis$y$pc_milk_yield[i])
  lessReplace_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
}

lessReplace_90_farms<-lessReplace_diagnosis$x$Scenario[which(lessReplace_diagnosis$y$confidence_level>0.9)]

lessReplace_diagnosis$x["Farm"]<-sapply(lessReplace_diagnosis$x$Scenario,
                                  function(x) as.numeric(strsplit(x,"_")[[1]][2]))

lessReplace_structures<-merge(lessReplace_diagnosis$x,farm_structures)
lessReplace_structures<-list(x=lessReplace_structures, y=lessReplace_diagnosis$y)

to_plot<-cbind(Farm=lessReplace_structures$x$Farm,Confidence_level=lessReplace_structures$y$confidence_level,
               lessReplace_structures$x[,c(paste0("breed_",1:8),"unproductive_bulls","replacement_males")])
to_plot$other_animals<-1-to_plot$replacement_males

melted<-melt(to_plot[,c("Farm","Confidence_level","other_animals","replacement_males")], id.vars=c("Farm","Confidence_level"))
melted<-melted[which(!is.na(melted$value)),]

melted$Confidence_level_rounded<-round(to_plot$Confidence_level,2)*100

ggplot(melted, aes(group=variable, fill=variable, y=value , x=Confidence_level_rounded)) + 
    geom_bar(position="fill", stat="identity") + coord_flip() +
  labs(y="Share of animal type in the herd",
       x="Probability of lower emissions intensity (%)",
       fill="Animal type") + scale_fill_manual(values= c("replacement_males" = "burlywood",
                                                         "other_animals" = "darkslategrey"),
                                               labels = c("Replacement males","Other animals")) +
  theme_bw()

ggsave("Fig_18_lessReplace_diagnostics.png",width=5,height=5)
